An Analysis of Health & Lifestyle Factors Associated with Sleep Quality

### 2024-05-03 products/manuscript/Manuscript.qmd
The End
Finally done!

For our project, we obtained a dataset from Kaggle.com containing 373
observations of 13 variables from a study on different factors that may
affect sleep quality. Variables of interest from the study included
physical activity level, age, sex, occupation, stress, gender, BMI,
blood pressure, and sleep duration. In
[RStudio](https://posit.co/download/rstudio-desktop/) (version 4.3.1),
we used a combination of packages to create models suitable for
answering our primary question: which variables are most important for
determining an individual's sleep quality and what effect do they have [@noautho
r_rstudio_nodate]?
After conducting an exploratory data analysis (EDA) we hypothesized that
stress, sleep duration, and occupation would have the largest impact on
sleep quality. We believed that stress and sleep quality would be
negatively correlated, sleep duration and quality would be positively
correlated, and that jobs with longer or more variable hours would
negatively affect sleep quality.

@@

The dataset originally contained 373 observations for 13 variables
covering a broad spectrum of health and lifestyle variables associated
with sleep quality. The curators of the dataset did not provide
information on how or where this data was collected, so we solely relied
on the codebook for variable definitions. We have added this codebook to
the supplementary materials file for ease of access but you can also
find it on the Kaggle page for this dataset. There were no missing
values or erroneous variables in the dataset that we needed to remove.
The "Blood.Pressure" variable was transformed to reflect the individual
systolic and diastolic blood pressure measurements from each subject. We
then chose to add an additional variable, labeled 'cat_bp' to reflect
the categorical blood pressure status of a subject based on the
[American Heart Association
guidelines](https://www.heart.org/en/health-topics/high-blood-pressure) [@noauthor_understanding_nodate].
Additionally we created a variable named 'StepsGroup' to determine the
categorical activity level of subjects in the dataset based on these
guidelines provided by [10,000
steps](https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps
/#:~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,
day%205%20Highly%20active%20is%20more%20than%2012%2C500) [@noauthor_counting_nodate].
Finally, we created an additional categorical variable called
'PhysicalActivityGroup' to categorically represent the level of physical
activity reported by a subject. The levels are differentiated by 30
minute intervals of weekly physical activity. The outcome of interest is
noted as "Quality.of.Sleep" that we will reference as Sleep Quality or
Quality of Sleep throughout this report.

>>>>

For our project, we obtained a dataset from Kaggle.com containing 373 observations of 13 variables from a study on different factors that may affect sleep quality. Variables of interest from the study included physical activity level, age, sex, occupation, stress, gender, BMI, blood pressure, and sleep duration. In [RStudio](https://posit.co/download/rstudio-desktop/) (version 4.3.1), we used a combination of packages to create models suitable for answering our primary question: which variables are most important for determining an individual's sleep quality and what effect do they have? After conducting an exploratory data analysis (EDA) we hypothesized that stress, sleep duration, and occupation would have the largest impact on sleep quality. We believed that stress and sleep quality would be negatively correlated, sleep duration and quality would be positively correlated, and that jobs with longer or more variable hours would negatively affect sleep quality.

@@

The dataset originally contained 373 observations for 13 variables covering a broad spectrum of health and lifestyle variables associated with sleep quality. The curators of the dataset did not provide information on how or where this data was collected, so we solely relied on the codebook for variable definitions. We have added this codebook to the supplementary materials file for ease of access but you can also find it on the Kaggle page for this dataset. There were no missing values or erroneous variables in the dataset that we needed to remove. The "Blood.Pressure" variable was transformed to reflect the individual systolic and diastolic blood pressure measurements from each subject. We then chose to add an additional variable, labeled 'cat_bp' to reflect the categorical blood pressure status of a subject based on the [American Heart Association guidelines](https://www.heart.org/en/health-topics/high-blood-pressure) [@noauthor_understanding_nodate]. Additionally we created a variable named 'StepsGroup' to determine the categorical activity level of subjects in the dataset based on these guidelines provided by [10,000 steps](https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps%20/#:~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,%20day%205%20Highly%20active%20is%20more%20than%2012%2C500) [@noauthor_counting_nodate]. Finally, we created an additional categorical variable called 'PhysicalActivityGroup' to categorically represent the level of physical activity reported by a subject. The levels are differentiated by 30 minute intervals of weekly physical activity. The outcome of interest is noted as "Quality.of.Sleep" that we will reference as Sleep Quality or Quality of Sleep throughout this report.
